---
title: "Assignment Part 1: Lake Ice"
author: "Quinn Thomas"
date: "2023-03-15"
output: html_document
---

```{r message=FALSE}
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
```

## Objective

Apply the tidymodel approach above to the ice-off data from the second module. Your goal is to translate the linear regression from the the lake ice module into the tidymodels machine learning framework.  

## Tips

- You should be using the same model and engine as in the ma
- Since the data are already processed you won't have any steps in your recipe.  Your recipe will just have a formula in the `recipe()`

## Step 1: Obtain data

First, copy data from ice lake phenology model into your data subdirectory in this module.  

**Question 1:** Provide code for reading in data and filtering the data to only be Lake Sunapee. 

**Answer 1:**

```{r}


```

## Step 2: Pre-process data

### Split data into training/testing sets

**Question 2:** Provide code for splitting data

**Answer 2:**

```{r}


```


#### Feature engineering using a recipe

**Question 3:** Provide code that defines the recipe for feature engineering

**Answer 3:**

```{r}


```


### Specify model and workflow

#### Define model type and mode

**Question 4:** Provide code that defines the model

**Answer 4:**

```{r}


```

#### Define workflow

**Question 5:** Provide code that defines the workflow

**Answer 5:**

```{r}


```

### Train model on Training Data

**Question 6:** Provide code that trains the model

**Answer 6:**

```{r}


```


### Predict Test Data

**Question 7:** Provide code that predicts the test data

**Answer 7:**

```{r}


```

### Evaluate model

**Question 8:** Provide code calculates the rmse and r2 for the test data

**Answer 8:**

```{r}


```

### Deploy model

#### Obtain new data

```{r}
new_data <- tibble(year = seq(2024, 2050, by = 1))
```

#### Make new prediction

**Question 9:**

Provide code that predicts the ice-off day of year for the new data

**Answer 9:**

```{r}


```

**Question 10:**

Plot predictions of the new data

**Answer 10:**

```{r}


```


