---
title: "Assignment Part 1: Lake Ice"
author: "Quinn Thomas"
date: "`r Sys.time()`"
output: github_document
---

```{r message=FALSE}
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
set.seed(100)
```

## Objective

Apply the tidymodel approach above to the ice-off data from the second module. Your goal is to translate the linear regression from the the lake ice module into the tidymodels machine learning framework.  

## Tips

- You should be using the same model and engine as in `machine-learning-101.Rmd`
- Since the data are already processed you won't have any steps in your recipe.  Your recipe will just have a formula in the `recipe()`

## Step 1: Obtain data

First, copy data from ice lake phenology model into your data subdirectory in this module.  

**Question 1:** Provide code for reading in data and filtering the data to only be Lake Sunapee. 

**Answer 1:**

```{r}


```

## Step 2: Pre-process data

### Split data into training/testing sets

**Question 2:** Provide code for splitting data

**Answer 2:**

```{r}


```


### Feature engineering using a recipe

**Question 3:** Provide code that defines the recipe for feature engineering

**Answer 3:**

```{r}


```


## Step 3: Specify model and workflow

### Define model type and mode

**Question 4:** Provide code that defines the model

**Answer 4:**

```{r}


```

### Define workflow

**Question 5:** Provide code that defines the workflow

**Answer 5:**

```{r}


```

## Step 4: Train model on Training Data

**Question 6:** Provide code that trains the model

**Answer 6:**

```{r}


```


## Step 5: Predict Test Data

**Question 7:** Provide code that predicts the test data

**Answer 7:**

```{r}


```

## Step 6: Evaluate model

**Question 8:** Provide code calculates the rmse and r2 for the test data

**Answer 8:**

```{r}


```

## Step 7:  Deploy model

### Obtain new data

```{r}
new_data <- tibble(year = seq(2024, 2050, by = 1))
```

### Make new prediction

**Question 9:** Provide code that predicts the ice-off day of year for the new data

**Answer 9:**

```{r}


```

**Question 10:** Plot the predictions of the new data (year vs. doy for the new data)

**Answer 10:**

```{r}


```

## Knitting and committing

Remember to Knit your document as a `github_document` and comment+push to GitHub your code, knitted document, and any files in the `figure-gfm` subdirectory that was created when you knitted the document.

