---
title: "Assignment Part 2: new ML model"
author: "YOUR NAME"
format:
  html:
    embed-resources: true
---

```{r message=FALSE}
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
set.seed(100)
```

## GitHub link

Provide link to your assignment GitHub repository here:

## Objective

Using machine learning and tidymodels to predict how much carbon is stored in vegetation across the U.S.

## Expections

This assignment requires the following

- Use two different models to generate predictions of plot-level forest carbon

- Your two models can't be both `linear_reg` or `rand_forest` (we have already used these).  You can use one or the other as one of your two models, just not both. One of your models must use hyper-parameter tuning.

- You must use more predictors than used in the `tutorial/tutorial1-machine-learning-101.qmd` and `tutorial/tutorial2-example-with-tuning.Rmd`.  You do not have to use all predictors.  In fact, one of the decisions that you will be making is which predictors you will include.  Do not include variables that are not meant to be predictors.

- You must upload your predictions of the new carbon stock data to Canvas as a CSV.

### Guidance

You will need to explore the different models/engines available in the tidy models package. The link below will show the list of available models. Clicking on a model type will show the different engines available to use.  

**[Alternative modeling approaches](https://parsnip.tidymodels.org/reference/index.html#models)**

Here is a list of models that may be particularly useful for the data in this assignment.

- [Single layer neural network](https://parsnip.tidymodels.org/reference/mlp.html)
- [Random forest](https://parsnip.tidymodels.org/reference/rand_forest.html)
- [Automatic Machine Learning](https://parsnip.tidymodels.org/reference/auto_ml.html)
- [Boosted trees](https://parsnip.tidymodels.org/reference/boost_tree.html)
- [Linear Regression](https://parsnip.tidymodels.org/reference/linear_reg.html)

The documentation of the engine will describe any pre-processesing that is required for the data and list the hyperparameters that are available for tuning. For example [here](https://parsnip.tidymodels.org/reference/details_rand_forest_ranger.html) is the documentation for the `ranger` engine within the `rand_forest` model

Once you start pre-processing your data using a recipe, the following link will show the different "steps" that you can use.  Clicking on the step will show how to use it in a recipe.  

**[Different recipes steps for feature engineering](https://recipes.tidymodels.org/reference/index.html)**

There are broad classes of recipe steps that you may find helpful for the data that you are working within this assignment.  You will not be using all the available steps.

- [Imputation (fill NA values)](https://recipes.tidymodels.org/reference/index.html#step-functions-imputation)
- [Transform variables](https://recipes.tidymodels.org/reference/index.html#step-functions-individual-transformations)
- [Add dummy variables](https://recipes.tidymodels.org/reference/index.html#step-functions-dummy-variables-and-encodings)
- [Create variable interaction](https://recipes.tidymodels.org/reference/index.html#step-functions-interactions)
- [Normalize variables (e.g., center on 0 and scale sd)](https://recipes.tidymodels.org/reference/index.html#step-functions-normalization)
- [Filter rows and select columns](https://recipes.tidymodels.org/reference/index.html#step-functions-filters)
- [Operate on rows](https://recipes.tidymodels.org/reference/index.html#step-functions-row-operations)

## Step 0: Warm-up

**Question 1:** What two model and engines did you select to use?

**Answer 1:**

Model 1:

Model 2:

**Question 2:** Provide a high level description of the modeling approaches you chose in Question 11 (max of two sentences each)

**Answer 2:**

Model 1:

Model 2:


**Question 3:** Do either models require any special data pre-processing? If so, describe the steps needed.

**Answer 3:**

Model 1:

Model 2:

**Question 4:** List any hyper-parameters that can be tuned in the models

Model 1:

Model 2: 

**Answer 4:**

## Step 1: Obtain data

```{r}
biomass_data <- read_csv("data/neon_biomass.csv", show_col_types = FALSE)
```

```
spc_tbl_ [1,000 × 17] (S3: spec_tbl_df/tbl_df/tbl/data.frame)
 $ plotID    : chr - identifier of plot
 $ daylength : num - mean annual daylength (s/day)
 $ precip    : num - mean annual precipitation (mm)
 $ range     : num - mean annual daily range in temperature (day max - day min; C)
 $ tavg      : num - mean annual temperature (C)
 $ solar     : num - mean annual solar radiation (W/m2)
 $ tmax      : num - mean annual daily max temperature (C)
 $ tmin      : num - mean annual daily min temperature (C)
 $ vpd       : num - mean annual vapor pressure deficient (Pa)
 $ elevation : num - elevation of plot (meters)
 $ nlcdClass : chr - land-cover classification of plot
 $ lat       : num - latitude of plot (degrees W)
 $ long      : num - longitude of plot (degrees E)
 $ plotType  : chr - type of plot (tower or distributed)
 $ ndvi      : num - normalized difference vegetation index of plot from MODIS (unitless)
 $ siteID    : chr - identifier of site that includes the plot
 $ plot_kgCm2: num - vegetation carbon stock of plot (kgC/m2)
```

## Step 2: Pre-process data

### Split data into training/testing sets

**Question 5:** Provide code for splitting data

**Answer 5:**

```{r}


```

### Split training data into folds

This step is for your model(s) and engine(s) that have hyper-parameters that can be tuned.

**Question 6:** Provide code for splitting data into folds. If your model does not require parameter tuning, then please just go ahead and state that as your answer rather than providing the code.

**Answer 6:**

```{r}


```

### Feature engineering using a recipe

**Question 7:** Provide code that defines the recipe for feature engineering. 

**Answer 7:**


```{r}


```

## Step 3: Specify model and workflow

### Define model type and mode

**Question 8:** Provide code that defines your two models (model type + engine).

**Answer 8:**

Model 1:

```{r}


```

Model 2:

```{r}


```

### Define workflow

**Question 9:** Provide code that defines the workflow

**Answer 9:**

Model 1:

```{r}


```

Model 2:

```{r}


```

## Step 4: Train model on Training Data

### Estimate best hyper-parameters using tuning

**Question 10:** Provide code that tunes the hyper-parameters of your model (for a model that has hyper-parameters, you are required tune at least one of them).  If your model+engine does not have hyper-parameters that can be tuned, then state that as your answer.

**Answer 10:**

Model 1:

```{r}


```

Model 2: 

```{r}


```

### Update workflow with best hyper-parameters

**Question 11:** Provide code that updates the workflow with the best hyper-parameters. If your model+engine does not require tuning, then state that as your answer.

**Answer 11:**

Model 1:

```{r}


```

Model 2: 

```{r}


```

### Fit to all training data

**Question 12:** Provide code that fits your model to the training data. Use the best hyper-parameter if you tuned them.

**Answer 12:**

Model 1: 

```{r}


```

Model 2:

```{r}


```

## Step 5: Predict Test Data

**Question 13:** Provide code that predicts the test data for each model

**Answer 13:**

Model 1: 

```{r}


```

Model 2: 

```{r}


```

## Step 6: Evaluate models

**Question 14:** Provide code that calculates the `rmse` and `r2` for the test data.  Make sure that the table with `rmse` and `r2` is shown in your rendered document.

**Answer 14:**

Model 1: 

```{r}


```

Model 2: 

```{r}


```

## Step 7: Retrain model

**Question 15:** Select one of your two models to proceed with generating prediction that you will use to predict new NEON data and contribute to the prediction challenge.  Why did you select that model?

**Answer 15:**

**Question 16:** Provide code that retrains your selected model on the full dataset.

**Answer 16:**

Model 1:
```{r}

```

Model 2:
```{r}


```

## Step 8: Deploy model

Use your selected model to predict new data. The new data is a data frame with all the same columns as the training and testing set above but with `NA` values for the biomass. I have the true biomass values and will calculate `rmse` for your predictions. Therefore you need to predict the biomass and upload your CSV to Canvas and I will download them as a set and run the analysis in class to examine the "scores"

```{r}
new_data <- read_csv("data/neon_biomass_new.csv", show_col_types = FALSE) 
```


**Question 17:** Provide code to predict the new data

**Answer 17:**

**Question 18:** Update the code below to write your submissions to a csv.

Be sure to change the `name` to your last name.

Be sure to change your model to the exact name of the model you used.  For example, we have used `rand_forest` and `linear_reg`.  Do not provide your engine.  We will use this column to understand how performance varies with modeling approach.

**Answer 18:**

```{r}
name <- "ADD YOUR LAST NAME HERE"
model_type <- "ADD YOUR MODEL TYPE"
new_predicted <- bind_cols(new_data, new_predictions) |>
  mutate(name = name,
         model = model_type) |>
  select(plotID, name, .pred)
write_csv(new_predicted, file = paste0(name,"-submission.csv"))
```

**Question 19:** Find your CSV on your computer from Question 18 and upload to Canvas.

**Answer 19:**

## Render and committing

Remember to Render your document as an `HTML` and comment+push to GitHub your code and rendered HTML document.  Use at least three commits that you push to GitHub. 

# Attribution

Include citation of any AI-generated assistance or discussion with classmates (per policy in syllabus). Proper documentation of AI-generated assistance includes the prompt, the source (e.g., ChatGPT), and the significant parts of the response.  Proper documentation of discussion with classmates include listing their names and the components discussed.  


